Synthesizing Symmetric Lenses

### 2018-05-29 papers/SymmetricLenses/main.tex
add text ahh

% begin introduction
\section{Introduction}
% state of the world
Similar data can come in varieties of formats: scheduled jobs can be stored as
cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
be provided to clients in legacy or modern formats; calendar events can be
stored in CSV files or ICS files. Oftentimes, we want to synchronize files
across formats: when I edit a task to run more often on my Linux machine, I want
this task to run more often on my Windows machine; when I update data through a
legacy API, I want those changes to be reflected when I retrieve it through a
modern API; when I create a new calendar event in my ICS file, I want my CSV
calendar files to be updated to reflect this change.

@@

A better approach is to synthesize such conversion functions. Recent research
has gone into Optician, a tool for automatically synthesizing such conversion
functions from input-output examples and formal descriptions of the data
formats~\cite{optician,quotient-optician}. However, this work assumes that all
data fields present in one format are present in the other, but this is not
always true. Oftentimes, fields present in one format are not present in the
other: the commands scheduled to run on Linux machines are typically different
than the commands scheduled to run on Windows machines; updated APIs can expose
new information, or omit legacy information; ICS files contain a URL field that
Google Calendar CSV files do not expose.

\emph{Symmetric lenses} formalize such conversion functions. However, these
symmetric lenses use external state for synchronization, in the form of a
\emph{complement}. This makes them difficult to use in real-world
synchronization tasks, as we must maintain the files we synchronize as well as
maintaining the complements. The complement further complicates synthesis, as
input-output examples are no longer sufficient, instead edit sequences must be
provided.

@@

We enhance Optician with simple symmetric lens synthesis capabilities. This
extension synthesizes symmetric lenses from a suite of input-output examples and
a pair of regular expressions describing the data formats. The synthesis
algorithm uses type-directed synthesis on the two regular expressions to
synthesize a lens that is well-typed between the provided regular expressions
and satisfies the provided input-output examples.

@@

We operate under the assumption that users wish to synchronize as much data as
possible, and try to synchronize as much data as the types and examples allow.
This task gets harder when there are more complex decisions involved. For
example, if I have two fields in one format, and one in the other, which field
on the left should be synchronized to the one of the right? Or would it be
better for the field on the right to be synchronized to a mish-mash of
subcomponents from each field? To answer these questions, we develop a cost
metric for reasoning about which lens synchronizes the most information.

@@

Our second algorithm, \GreedySynth, utilizes alternative languages for
stochastic regular expressions and lenses, \emph{stochastic DNF regular
  expressions} and \emph{symmetric DNF lenses}. While these languages are hard
to read and understand, they permit an efficient type-directed synthesis
algorithm. We develop our cost metric on DNF lenses, and provide a greedy
algorithm that tries to find a symmetric DNF lens of minimal cost.

@@

In summary, our contributions are:
\begin{enumerate}
\item We develop \emph{simple symmetric lenses} a new formulation for
  bidirectional programs that can project data from both sides~(\S\ref{sec:ssl}),
  and describe a language that expresses such lenses on string
  data~(\S\ref{sec:ssl-string}).
\item We develop a means to traverse stochastic regular expression equivalences,
  while preserving probability distributions and
  languages~(\S\ref{sec:stoch-rx}).
\item We develop stochastic DNF regular expressions and DNF lenses. We develop a
  sound algorithm for converting stochastic regular expressions to stochastic
  DNF regular expressions~(\S\ref{sec:dnf-regex-lens}), and a
  information-theoretic cost metric on DNF
  lenses~(\S\ref{sec:symmetric-dnf-lenses}).
\item We extended Boomerang with our simple symmetric lens combinators. We
  updated the Optician synthesis system to synthesize symmetric lenses with our
  new algorithm.
\item We evaluate our synthesis algorithm on Opticians benchmark suite as well
  as new, symmetric benchmarks, and find that we lose very little synthesis
  time, while being significantly more expressive (\S\ref{sec:evaluation}).
\end{enumerate}
% end introduction

@@

To synchronize these files, a few functions are needed. Two \emph{creates} are
needed; if a crontab is initialized, a synchronized Scheduled Tasks file is also
created, and vice versa. Furthermore, we require \emph{puts}, functions that
take an updated crontab and an old Scheduled Tasks file, and output a new
Scheduled Tasks file with the information synchronized to the updated crontab.
To have a simple symmetric lens is to have these required synchronization
functions. Formally, a simple symmetric lens between $X$ and $Y$ comprises 4
functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a lens, one must satisfy the
following ``round-tripping'' laws that guarantee identity edits won't change the
file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

@@

Prior work on lens synthesis works with two communicating synthesizers, one that
proposes regular expressions, semantically but not syntactically equivalent to
each other, and another that finds lenses between them, if there are any lenses
between them that don't use a type equivalence rule involving star equivalences.
We utilize this same framework, except instead of having the second synthesizer
find \emph{any} lens between the types, \GreedySynth instead uses a greedy
algorithm to try finding the \emph{best} lens between the types.

@@

\section{Simple Symmetric Lenses}
\label{sec:ssl}
Intuitively, simple symmetric lenses are much like symmetric lenses, but they
avoid complements by keeping all information required to stitch edits to a data
source in the data source itself. Instead of having puts operate on a
complement, they operate on both data sources, where the alterations of the
first argument update the second. Furthermore, on initial creations of the
files, instead of using an $init$ complement, two additional functions,
$\CreateR$ and $\CreateL$ are used.

>>>>

% begin introduction
\section{Introduction}
% state of the world
Similar data can come in varieties of formats: scheduled jobs can be stored as
cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
be provided to clients in legacy or modern formats; calendar events can be
stored in CSV files or ICS files. Oftentimes, files should be synchronized
across these formats: when I edit a task to run more often on my Linux machine,
I want this task to run more often on my Windows machine; when I update data
through a legacy API, I want those changes to be reflected when I retrieve it
through a modern API; when I create a new calendar event in my ICS file, I want
my CSV calendar files to be updated to reflect this change.

@@

A better approach is to synthesize such conversion functions. Prior work on
synthesizing these functions assumes that all data fields present in one format
are present in the other, but this is not always true. Oftentimes, fields
present in one format are not present in the other: the commands scheduled to
run on Linux machines are typically different than the commands scheduled to run
on Windows machines; updated APIs can expose new information, or omit legacy
information; ICS files contain a URL field that Google Calendar CSV files do not
expose.

\emph{Symmetric lenses} formalize such conversion functions. However, these
symmetric lenses use external state for synchronization, in the form of a
\emph{complement}. This makes them difficult to use in real-world
synchronization tasks, as users must maintain the synchronized files synchronize
as well as the complements. The complement further complicates synthesis, as
input-output examples are no longer sufficient, instead edit sequences must be
provided.

@@

We provide a tool for synthesizing simple symmetric lenses. This tool
synthesizes symmetric lenses from a suite of input-output examples and a pair of
regular expressions describing the data formats. The synthesis algorithm uses
type-directed synthesis on the two regular expressions to synthesize a lens that
is well-typed between the provided regular expressions and satisfies the
provided input-output examples.

@@

Operate under the assumption that users wish to synchronize as much data as
possible, our algorithm try to synchronize as much data as the types and
examples allow. This task gets harder when there are more complex decisions
involved. For example, if I have two fields in one format, and one in the other,
which field on the left should be synchronized to the one of the right? Or would
it be better for the field on the right to be synchronized to a mish-mash of
subcomponents from each field? To answer these questions, we develop a cost
metric for reasoning about which lens synchronizes the most information.

@@

Our second algorithm, \GreedySynth, utilizes alternative languages for
stochastic regular expressions and lenses, \emph{stochastic DNF regular
  expressions} and \emph{symmetric DNF lenses}. While these languages are hard
to read and understand, they permit an efficient type-directed synthesis
algorithm. We develop our cost metric on DNF lenses, and provide a greedy
algorithm that tries to find a symmetric DNF lens of minimal cost. The
synthesized symmetric DNF lens is then converted to a lens in the original
simple symmetric lens language.

@@

In summary, our contributions are:
\begin{enumerate}
\item We develop \emph{simple symmetric lenses} a new formulation for
  bidirectional programs that can project data from both sides~(\S\ref{sec:overview}),
  and describe a language that expresses such lenses on string
  data~(\S\ref{sec:ssl}).
\item We develop a means to traverse stochastic regular expression equivalences,
  while preserving probability distributions and
  languages~(\S\ref{sec:stoch-rx}).
\item We develop stochastic DNF regular expressions and DNF lenses. We develop a
  sound algorithm for converting stochastic regular expressions to stochastic
  DNF regular expressions~(\S\ref{sec:dnf-regex-lens}), and a
  information-theoretic cost metric on DNF
  lenses~(\S\ref{sec:symmetric-dnf-lenses}).
\item We extended Boomerang with our simple symmetric lens combinators. We
  updated the Optician synthesis system to synthesize symmetric lenses with our
  new algorithm.
\item We evaluate our synthesis algorithm on Opticians benchmark suite as well
  as new, symmetric benchmarks, and find that we lose very little synthesis
  time, while being significantly more expressive (\S\ref{sec:evaluation}).
\end{enumerate}
% end introduction

@@

To synchronize these files, a few functions are needed. Two \emph{creates} are
needed; if a crontab is initialized, a synchronized Scheduled Tasks file is also
created, and vice versa. Furthermore, \emph{puts} are needed, functions that
take an updated crontab and an old Scheduled Tasks file, and output a new
Scheduled Tasks file with the information synchronized to the updated crontab.
To have a simple symmetric lens is to have these required synchronization
functions. Formally, a simple symmetric lens between $X$ and $Y$ comprises 4
functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a lens, one must satisfy the
following ``round-tripping'' laws that guarantee identity edits won't change the
file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

@@

Prior work on lens synthesis works with two communicating synthesizers, one that
proposes regular expressions, semantically but not syntactically equivalent to
each other, and another that finds lenses between them, if there are any lenses
between them that don't use a type equivalence rule involving star equivalences.
We utilize this same framework, with two modifications. (1) The first
synthesizer does not propose regular expressions, but stochastic regular
expressions, equivalent to the originals. We prove that these proposed
stochastic regular expressions don't merely have the same languages as the
originals, but also the same probability distributions. (2) Instead of having
the second synthesizer find \emph{any} lens between the types, \GreedySynth
instead uses a greedy algorithm to try finding the \emph{best} lens between the
types. Furthermore, the first synthesizer proposes stochastic regular
expressions with equivalent languages and probability distributions, instead of
just equivalent languages.

@@

\section{Simple Symmetric Lenses}
\label{sec:ssl2}
Intuitively, simple symmetric lenses are much like symmetric lenses, but they
avoid complements by keeping all information required to stitch edits to a data
source in the data source itself. Instead of having puts operate on a
complement, they operate on both data sources, where the alterations of the
first argument update the second. Furthermore, on initial creations of the
files, instead of using an $init$ complement, two additional functions,
$\CreateR$ and $\CreateL$ are used.
